---
author: "Ryan"
execute:
    echo: FALSE
    message: FALSE
    warning: FALSE
---
## Distrubution of Bachelor Degree In Los Angeles 

This interactive map displays the **total population** by **census tract** in **Los Angeles County**, based on data from the **2020 American Community Survey (ACS)**. Each tract is shaded in varying intensities of **blue**, where **darker shades** represent areas with **higher population counts**. Hovering over a tract reveals its unique **GEOID** and **population estimate**. This visualization helps highlight **population density patterns** across the county, providing insight into how people are distributed **geographically**. The use of **tidycensus** and **leaflet** enables a powerful combination of **demographic data access** and **web-based mapping**.


```{r}
library(tidycensus)
library(tidyr)
library(primer.data)
library(viridis)
library(leaflet)
library(dplyr)
library(sf)

```
```{r}

la_tracts <- get_acs(
  geography = "tract",
  variables = "B01003_001",  # Total population
  state = "CA",
  county = "Los Angeles",
  year = 2020,
  geometry = TRUE,
  progress_bar = FALSE
)
```


```{r}
library(tidycensus)

# You already have this:
# la_tracts <- get_acs(...)

# Step 1: Prepare a color palette
pal <- colorNumeric(
  palette = "Blues",
  domain = la_tracts$estimate,
  na.color = "transparent"
)

# Step 2: Create informative labels
labels <- sprintf(
  "<strong>Tract:</strong> %s<br/><strong>Population:</strong> %s",
  la_tracts$GEOID,
  formatC(la_tracts$estimate, format = "d", big.mark = ",")
) %>% lapply(htmltools::HTML)

# Step 3: Build the leaflet map
leaflet(data = la_tracts) %>%
  addProviderTiles("CartoDB.Positron") %>%  # Clean, modern basemap
  addPolygons(
    fillColor = ~pal(estimate),
    weight = 0.3,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 1.5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE
    ),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "13px",
      direction = "auto"
    )
  ) %>%
  addLegend(
    pal = pal,
    values = ~estimate,
    opacity = 0.7,
    title = "Total Population",
    position = "bottomright"
  )

```